---
title: "Peer Assignment 2"
output: html_document
---

This R Markdown document was created for Peer Assignment 2 for the COursera class Reproducible Research.


# Synopsis



## Processing the Data

The first step is to read the data ("repdata-data-StormData.csv") into R. The following code will acomplish this. Make sure that the data exists and is unzipped into your working directory. Once we read the data, we will also view the first few rows to get a sense of how the data is structured and make sure that it was read properly.

```{r}
weather_data <- read.table("repdata-data-StormData.csv", header=TRUE, sep=",")
head(weather_data)
```

Additional data processing is described where appropriate in the Results section below. 

#Results

##Most Hardful Weather-Related Events

The first question we need to answer using these data is:

1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

To answer this question we must utilize the EVTYPE variable to determine the number of fatailities and injuries per event type.

First we will aggregate the total number of injuries by event type, choosing the top 10.

```{r}
total_injuries <- aggregate(weather_data$INJURIES, list(Event = weather_data$EVTYPE), sum)
top_total_injuries <- total_injuries[order(total_injuries$x, decreasing=TRUE), ]
top_10_total_injuries <- head(top_total_injuries, 10)
top_10_total_injuries
```

Do the same for The total number of fatalities by event type, choosing the top 10. 

```{r}
total_fatalities <- aggregate(weather_data$FATALITIES, list(Event = weather_data$EVTYPE), sum)
top_total_fatalities <- total_fatalities[order(total_fatalities$x, decreasing=TRUE), ]
top_10_total_fatalities <- head(top_total_fatalities, 10)
top_10_total_fatalities
```

Now, having found the top 10 most dangerous weather events (by injuries and fatalities) we can make two bar charts to better visualize the data. 

```{r}
par(mfrow=c(1,2))
barplot(top_10_total_injuries$x, names.arg=top_10_total_injuries$Event, cex.names=0.45, las=3, main="Top 10 Events by Total Injuries", ylab="Total Number of Weather-Related Injuries", col="blue")
barplot(top_10_total_fatalities$x, names.arg=top_10_total_fatalities$Event, cex.names=0.45, las=3, main="Top 10 Events by Total Fatalities", ylab="Total Number of Weather-Related Fatalities", col="red")
```

Referencing these three charts we can see that the top 5 most harmful weather-related events are:

1) Tornado
2) Heat
3) Flood
4) Lightning 
5) wind-related Events

##Weather-Related Events with the Greatest Economic Consequences

The second question that we need to answer is:

2) Across the United States, which types of events have the greatest economic consequences?

We can determine the level of economic impact for various weather-related events by examining the variables related to property damage (PROPDMG) as well as crop damage (CROPDMG) within the original dataset. Additional variables used were PROPDMGEXP and CROPDMBGEXP which include exponent of the value in the preceeding variable. For example, "K" correspondes to 1000.

First we need to clean the exponent values and replace them with the proper value. Some values had characters and other were lower or upper case. 

```{r}
#Subsitute CROPDMBGEXP with corresponding numerical value



#Subsitute PROPDMGEXP with corresponding numerical value
weather_data$PROPDMGEXP <- sub("^$|[-?+]", "0",weather_data$PROPDMGEXP)
weather_data$PROPDMGEXP <- sub("[Hh]", "100",weather_data$PROPDMGEXP)
weather_data$PROPDMGEXP <- sub("[Kk]", "1000",weather_data$PROPDMGEXP)
weather_data$PROPDMGEXP <- sub("[Mm]", "1000000",weather_data$PROPDMGEXP)
weather_data$PROPDMGEXP <- sub("[Bb]", "1000000000",weather_data$PROPDMGEXP)
```



#Conclusion